Left-linear rules in adhesive categories

### 2024-02-14 icalp/fusione.tex
minori

  \item $\mor(\X)$, $\mon(\X)$ and $\reg(\X)$ will denote the class of
    all arrows, monos and regular monos of $\X$, respectively.

  \item Given an integer $n\in \mathbb{Z}$, $[0,n]$ will denote the
    set of natural numbers less than or equal to $n$, assuming that
    $[0,n]=\emptyset$ whenever $n<0$.
  \end{itemize}
\end{notation}

@@

    We will call a $\mor(\X)$-Van Kampen ($\mor(\X)$-stable) square simply a \emph{Van Kampen} (\emph{stable}) square.
  \end{definition}}
\parbox{2cm}{$\xymatrix@C=10pt@R=10pt{&A'\ar[dd]|\hole_(.65){a}\ar[rr]^{f'} \ar[dl]_{m'} && B' \ar[dd]^{b} \ar[dl]_{n'} \\ C'  \ar[dd]_{c}\ar[rr]^(.7){g'} & & D' \ar[dd]_(.3){d}\\&A\ar[rr]|\hole^(.65){f} \ar[dl]^{m} && B \ar[dl]^{n} \\C \ar[rr]_{g} & & D \\A \ar[dd]_{m}\ar[rr]^{f}&&B\ar[dd]^{n}\\ \\C\ar[rr]_{g} &&D}$ }

@@

We will write $m\colon X\mto Y$ to denote that an arrow $m\colon X\to Y$ belongs to $\mathcal{M}$.

@@

\begin{example}
  \label{ex:adhesive}
  Every topos is adhesive~\cite{lack2006toposes}, thus in particular
  $\cat{Set}$ is adhesive. By the closure properties recalled before,
  every presheaf $[\cat{X},\cat{Set}]$ is adhesive and in particular
  the category $\cat{Graph} = [ E \rightrightarrows V, \cat{Set}]$ is
  adhesive where $E \rightrightarrows {V}$ is the two object category
  with two morphisms $s,t \colon{E} \to {V}$. In a similar way,
  various categories of hypergraphs can be shown to be adhesive, 
  such as term graphs and hierarchical graphs~\cite{CastelnovoGM24},
  as well as the category $\cat{sGraphs}$ of simple graphs, i.e. graphs without
  parallel edges, is $\reg{(\cat{sGraphs})}$-adhesive~\cite{xxx}
.%\todo{Add reference for simple graphs, add
 %   hierarchical graphs and term
 %   graphs\cite{castelnovo2023thesis,CastelnovoGM22}).}
\end{example}

@@

  \vspace{.1cm}
  \noindent
  \parbox{10cm}{\hspace{15pt}Given two objects $G$ and $H$ and a rule
    $\rho=(l,r)$ in $\R$, a \emph{direct derivation $\mathscr{D}$ from
      $G$ to $H$ applying the rule $\rho$} is a diagram as the one
    aside, in which both squares are pushouts. The arrow $m$ is called
    the \emph{match} of the derivation, while $h$ is its
    \emph{back-match}.  We will denote a direct derivation $\dder{D}$
    between $G$ and $H$ as $\dder{D}\colon G\Mapsto H$. }
  \parbox{3cm}{$\xymatrix{L \ar[d]_{m}& K \ar[d]^{k}\ar@{>->}[l]_{l}
      \ar[r]^{r} & R \ar[d]^{h}\\G & \ar@{>->}[l]^{f} D \ar[r]_{g}&
      H}$}
\end{definition}

@@

\begin{definition}[Derivation]
  Let  $\X$ be an $\mathcal{M}$-adhesive category and 
  $(\X, \R)$ a left-linear rewriting system. 
  Given $G, H \in \X$, a \emph{derivation}
  $\der{D}$ with source $G$ and target $H$, written
  $\der{D}: G \Mapsto H$, is defined as a sequence
  $\{\dder{D}_i\}_{i=0}^n$ of direct derivations such that
  \begin{enumerate}
  \item $\dder{D}_i : G_i \Mapsto G_{i+1}$ is a direct derivation for
    every index $i$;
  \item $G_0=G$ and $G_{n+1}=H$.
  \end{enumerate}
  Moreover, we have an \emph{empty derivation} $G : G \Mapsto G$ for each $G \in \X$.
  %
  We will denote by $\lgh(\der{D})$ the \emph{length} of derivation $\der{D}$.
 %
  \full{We will call the number $n+1$ the \emph{length} of the derivation,
  denoted by $\lgh(\der{D})$. We will also say that an empty
  derivation has length $0$.}

@@

\noindent
\parbox{8cm}{
  \begin{definition}
    Let  $\X$ be an $\mathcal{M}$-adhesive category and 
    $(\X, \R)$ a left-linear rewriting system. 
    An \emph{abstraction equivalence} between two derivations $\der{D}$
    and $\der{D'}$ with the same length and
    $r(\der{D})=r(\der{D}')$, is a family of isomorphisms
    $\{\phi_X\}_{X\in \{G_i,D_i\}}$ such that the diagram on the right commutes
    for every $i\in [0, \lgh(\der{D})]$. We
    will say that $\der{D}$ are $\der{D}'$ are \emph{abstraction
      equivalent}, and use the symbol $\der{D}\equiv_a \der{D}'$, if there exists an abstraction equivalence between
    them.
  \end{definition}
}
\parbox{4cm}{\vspace{-1.35em}$\xymatrix@C=40pt{G'_i&D'_i \ar[r]^{g'_i}
    \ar@{>->}[l]_{f'_i}&G'_{i+1}\\ L_i \ar[u]^{m'_i} \ar[d]_{m_i}& K_i
    \ar[u]^{k'_i} \ar[d]_{k_i} \ar[r]^{r_i} \ar@{>->}[l]_{l_i}
    &R_i \ar[u]^{h'_i} \ar[d]_{h_i}\\G_i
    \ar@/_.45cm/[uu]_(.35){\phi_{G_i}}|\hole&D_i\ar@{>->}[l]^{f_i}\ar@/_.45cm/[uu]_(.35){\phi_{D_i}}|\hole
    \ar[r]_{g_i}&G_{i+1}\ar@/_.45cm/[uu]_{\phi_{G_{i+1}}}}$}

@@

\begin{proof}
  Let us start fixing some notation. Given $X_1\in \Delta(\der{D}_1)$,
  $X_2\in \Delta(\der{D}_2)$, $X'_1\in \Delta(\der{D}'_1)$,
  $X'_2\in \Delta(\der{D}'_2)$, we will denote their coprojections
  into, respectively, $\lpro \der{D}_1\rpro$, $\lpro \der{D}'_2\rpro$,
  $\lpro \der{D}'_1\rpro$, $\lpro \der{D}'_2\rpro$ by
  $\iota_{1, X_1}\colon X_1\to \lpro\der{D}_1\rpro$,
  $\iota_{1, X_2}\colon X_2\to \lpro\der{D}_1\rpro$,
  $\iota'_{1, X'_1}\colon X'_1\to \lpro\der{D}'_1\rpro$ and
  $\iota'_{2, X'_2}\colon X'_2\to \lpro\der{D}_1\rpro$. Similarly,
  given $X\in \Delta(\der{D})$ and $\X'\in \Delta(\der{D}')$ we will
  use $\iota_X\colon X\to \tpro{D}$ and
  $\iota'_{X'}\colon X'\to \lpro \der{D}' \rpro$ for the
  coprojections.

  Now that these preliminaries matters have been addressed, we will proceed with the rest of the proof.. We split the proof in three cases.
  \begin{itemize}
  \item $\lgh(\der{D}_1)=0$. Thus $\lgh(\der{D}'_1)=0$ too  and we have
    \begin{align*}
      (\der{D}, \alpha, \omega) & =(\der{D}_2, \alpha_2\circ \omega_1^{-1}\circ \alpha_1, \omega_2) \\ (\der{D}', \alpha', \omega')&=(\der{D}'_2, \alpha'_2\circ (\omega'_1)^{-1}\circ \alpha'_1, \omega'_2)
    \end{align*}
    Moreover, in this case $\sigma$ must be $\id{\emptyset}$ and $\sigma+\tau$ must be equal to $\tau$. The thesis now follows from the commutativity of the following diagram.
    \[\xymatrix@C=35pt{& G_{1,0}\ar[dd]^{\xi_{\id{\emptyset}}} \ar[dr]^{\omega^{-1}_1}&&G_{2,0}\ar[r]^-{\iota_{2, G_{2,0}}} & \lpro \der{D}_2\rpro \ar[dd]^{\xi_{\tau}}\\\pi(G_{1,0})  \ar[ur]^{\alpha_1} \ar[dr]_{\alpha'_1}&& \pi(G_{2,0}) \ar[ur]^{\alpha_2} \ar[dr]_{\alpha'_2}\\& G'_{1,0} \ar[ur]_{(\omega'_1)^{-1}}&&G'_{2,0} \ar[r]_-{\iota'_{2, G_{2,0}}}& \lpro \der{D}'_2\rpro}\]

@@

  If a switch of $\dder{D}_0$ and $\dder{D}_1$ exists we will say that
  they are \emph{switchable}.

@@

\begin{notation} We will denote any switch of two switchable direct derivations $\dder{D}_0$ and $\dder{D}_1$ along $(i_0, i_1)$ by $S_{i_0, i_1}(\dder{D}_0, \dder{D}_1)$. Moreover, we will use $S_{i_0,i_1}(\dder{D}_1)$ and  $S_{i_0,i_1}(\dder{D}_0)$ to denote its components, in such a way that \[S_{i_0, i_1}(\dder{D}_0, \dder{D}_1)=S_{i_0,i_1}(\dder{D}_1)\cdot S_{i_0,i_1}(\dder{D}_0)\]
\end{notation}

@@

\begin{remark}
  Clearly, a permutation $\perm$ is the identity if and only if
  $\inv{\perm}=\emptyset$.
\end{remark}

@@

We also need to introduce the notion of \emph{switching sequences}.

@@

\begin{definition}[Switch equivalence]
  \label{de:switch-equivalence}
  Let $(\X, \R)$ be a left-linear rewriting system.  Let also
  $\der{D}, \der{E} : G \Mapsto H$ be derivations with the same
  length, $\der{D}=\{\dder{D}_{i}\}_{i=0}^n$ and
  $\der{E}=\{\dder{E}_{i}\}_{i=0}^n$. If
  $\dder{D}_i \cdot \dder{D}_{i+1}$ is a switch of
  $\dder{E}_i \cdot \dder{E}_{i+1}$ for some $i \in [0,n-1]$ \todo{A: ma manca ``and  $\dder{D}_j = \dder{E}_j$ for each $j \not \in \{i,i+1\}$''??} we write
  $\der{D} \shift{\transp{i}} \der{E}$.

  A \emph{switching sequence} is  a sequence $\{\der{D}_{k}\}_{k=0}^m$ of derivations such that 
  \[\der{D}_0 \shift{\transp{i_1}} \der{D}_1 \shift{\transp{i_2}} \ldots \shift{\transp{i_m}}
  \der{D}_m\]
 In such case, we will\todo{A: io cambierei tutti i ``we will'' in ``we''.} often denote by $\nu_{k}$ the transposition $\transp{i_k}$. Given $k \leq h$,
 we denote by $\nu_{k,h}$ the composition
 $\nu_h \circ \nu_{h-1} \circ \ldots \nu_k$. We will say that the
 switching sequence consists of inversions if for all
 $k \in \interval[0]{m}$ the transposition $\nu_k$ is an inversion for
 $\nu_{1,m}$.

  We will say that two derivations $\der{D}, \der{E}:G\Mapsto H$ are \emph{switch
    equivalent} if there is a switching sequence, possibly empty, $\{\der{D}_{k}\}_{k=0}^m$ such that
      \[\der{D}\equiv_a \der{D}_0 \shift{\transp{i_1}} \der{D}_1 \shift{\transp{i_2}} \ldots \shift{\transp{i_m}}
    \der{D}_m\equiv _a \der{E}\]
    We will denote this relation by $\der{D}\equiv^{sh} \der{E}$. 
\end{definition}

@@

  Consider the pullback of $g_0$ and $f_1$, getting $p_i : P \to D_i$
  for $i \in \{0,1\}$ and the mediating arrows $u_0\colon K_0\to P$ and
  $u_1\colon K_1\to P$ into the pullback object (see \Cref{prop:tec} for details). 
  % given by \Cref{prop:tec} .
  %
  We will say that $(i_0, i_1)$ is a \emph{strong independence pair} if
  the first two squares depicted  
  below are pushouts and if the pushout of $r_1 : K_1 \to R_1$ and $u_1 : K_1 \to P$ exists
  \[
    \xymatrix{
      K_0 \ar[r]^{r_0} \ar[d]_{u_0}& R_0 \ar[d]^{i_0} & K_1
      \ar@{>->}[r]^{l_1} \ar[d]_{u_1}& L_1 \ar[d]^{i_1}
       &K_1 \ar[r]^{r_1} \ar[d]_{u_1}& R_1\ar@{.>}[d]^{j_0}
      \\
      %
      P \ar[r]_{p_1} & D_1 &P \ar[r]_{p_0} & D_0
       & P \ar@{.>}[r]_{q_1} & Q_1
    }
  \]
\end{definition}

@@

\begin{definition}[Graph signature]
  Let $\mathcal{G}=(V, E, s, t)$ be a finite directed acyclic graph,
  we define the \emph{depth} of a vertex $v\in V$ as
  \[
    \dph(v)=\begin{cases}
      0 & \text{there is no $e\in E$ such that $t(e)=v$}\\
      m+1 &\text{where } m=\min\{\dph(s(e)) \mid e\in E \text{ such that } t(e)=v
      \} 
    \end{cases}
  \]
\end{definition}

@@

  Finally, in applications, one is interested in rewriting systems in which the right half of a rule is always mono but not regular in general, allowing us to identify more nodes. These rewring systems are all well-switching.

  \todo[inline]{Noi prendiamo sistemi di riscrittura left-linear e mono a destra, c'è una refernza?}  
\end{example}

@@

\subsection{Some results on $\mathcal{A}$-stable and $\mathcal{A}$-Van Kampen squares}
We will start proving some general results regarding $\mathcal{A}$-Van Kampen and $\mathcal{A}$-stable squares. Let us begin recalling some classical results about  about pullbacks and pushouts\cite{mac2013categories}.

@@

In this appendix we will provide the proofs of the main results of this paper.

>>>>

  \item $\mor(\X)$, $\mon(\X)$ and $\reg(\X)$ denote the class of
    all arrows, monos and regular monos of $\X$, respectively.

  \item Given an integer $n\in \mathbb{Z}$, $[0,n]$ denotes the
    set of natural numbers less than or equal to $n$, assuming that
    $[0,n]=\emptyset$ whenever $n<0$.
  \end{itemize}
\end{notation}

@@

    We call a $\mor(\X)$-Van Kampen ($\mor(\X)$-stable) square simply a \emph{Van Kampen} (\emph{stable}) square.
  \end{definition}}
\parbox{2cm}{$\xymatrix@C=10pt@R=10pt{&A'\ar[dd]|\hole_(.65){a}\ar[rr]^{f'} \ar[dl]_{m'} && B' \ar[dd]^{b} \ar[dl]_{n'} \\ C'  \ar[dd]_{c}\ar[rr]^(.7){g'} & & D' \ar[dd]_(.3){d}\\&A\ar[rr]|\hole^(.65){f} \ar[dl]^{m} && B \ar[dl]^{n} \\C \ar[rr]_{g} & & D \\A \ar[dd]_{m}\ar[rr]^{f}&&B\ar[dd]^{n}\\ \\C\ar[rr]_{g} &&D}$ }

@@

We write $m\colon X\mto Y$ to denote that an arrow $m\colon X\to Y$ belongs to $\mathcal{M}$.

@@

\begin{example}
  \label{ex:adhesive}
  Every topos is adhesive~\cite{lack2006toposes}, thus in particular
  $\cat{Set}$ is adhesive. By the closure properties recalled before,
  every presheaf $[\cat{X},\cat{Set}]$ is adhesive and in particular
  the category $\cat{Graph} = [ E \rightrightarrows V, \cat{Set}]$ is
  adhesive where $E \rightrightarrows {V}$ is the two object category
  with two morphisms $s,t \colon{E} \to {V}$. In a similar way,
  various categories of hypergraphs can be shown to be adhesive, 
  such as term graphs and hierarchical graphs~\cite{CastelnovoGM24}. We can also notice that the category $\cat{sGraphs}$ of simple graphs, i.e. graphs without
  parallel edges, is $\reg{(\cat{sGraphs})}$-adhesive~\cite{behr2022fundamentals} but not quasiadhesive.
\end{example}

@@

  \vspace{.1cm}
  \noindent
  \parbox{10cm}{\hspace{15pt}Given two objects $G$ and $H$ and a rule
    $\rho=(l,r)$ in $\R$, a \emph{direct derivation $\mathscr{D}$ from
      $G$ to $H$ applying the rule $\rho$} is a diagram as the one
    aside, in which both squares are pushouts. The arrow $m$ is called
    the \emph{match} of the derivation, while $h$ is its
    \emph{back-match}.  We denote a direct derivation $\dder{D}$
    between $G$ and $H$ as $\dder{D}\colon G\Mapsto H$. }
  \parbox{3cm}{$\xymatrix{L \ar[d]_{m}& K \ar[d]^{k}\ar@{>->}[l]_{l}
      \ar[r]^{r} & R \ar[d]^{h}\\G & \ar@{>->}[l]^{f} D \ar[r]_{g}&
      H}$}
\end{definition}

@@

\begin{definition}[Derivation]
  Let  $\X$ be an $\mathcal{M}$-adhesive category and 
  $(\X, \R)$ a left-linear rewriting system. 
  Given $G, H \in \X$, a \emph{derivation}
  $\der{D}$ with source $G$ and target $H$, written
  $\der{D}: G \Mapsto H$, is defined as a sequence
  $\{\dder{D}_i\}_{i=0}^n$ of direct derivations such that
  \begin{enumerate}
  \item $\dder{D}_i : G_i \Mapsto G_{i+1}$ is a direct derivation for
    every index $i$;
  \item $G_0=G$ and $G_{n+1}=H$.
  \end{enumerate}
  Moreover, we have an \emph{empty derivation} $G : G \Mapsto G$ for each $G \in \X$.
  %
  We denote by $\lgh(\der{D})$ the \emph{length} of derivation $\der{D}$.
 %
  \full{We will call the number $n+1$ the \emph{length} of the derivation,
  denoted by $\lgh(\der{D})$. We also say that an empty
  derivation has length $0$.}

@@

\noindent
\parbox{8cm}{
  \begin{definition}
    Let  $\X$ be an $\mathcal{M}$-adhesive category and 
    $(\X, \R)$ a left-linear rewriting system. 
    An \emph{abstraction equivalence} between two derivations $\der{D}$
    and $\der{D'}$ with the same length and
    $r(\der{D})=r(\der{D}')$, is a family of isomorphisms
    $\{\phi_X\}_{X\in \{G_i,D_i\}}$ such that the diagram on the right commutes
    for every $i\in [0, \lgh(\der{D})]$. We say that $\der{D}$ are $\der{D}'$ are \emph{abstraction
      equivalent}, and use the symbol $\der{D}\equiv_a \der{D}'$, if there exists an abstraction equivalence between
    them.
  \end{definition}
}
\parbox{4cm}{\vspace{-1.35em}$\xymatrix@C=40pt{G'_i&D'_i \ar[r]^{g'_i}
    \ar@{>->}[l]_{f'_i}&G'_{i+1}\\ L_i \ar[u]^{m'_i} \ar[d]_{m_i}& K_i
    \ar[u]^{k'_i} \ar[d]_{k_i} \ar[r]^{r_i} \ar@{>->}[l]_{l_i}
    &R_i \ar[u]^{h'_i} \ar[d]_{h_i}\\G_i
    \ar@/_.45cm/[uu]_(.35){\phi_{G_i}}|\hole&D_i\ar@{>->}[l]^{f_i}\ar@/_.45cm/[uu]_(.35){\phi_{D_i}}|\hole
    \ar[r]_{g_i}&G_{i+1}\ar@/_.45cm/[uu]_{\phi_{G_{i+1}}}}$}

@@

\begin{proof}
  Let us start fixing some notation. Given $X_1\in \Delta(\der{D}_1)$,
  $X_2\in \Delta(\der{D}_2)$, $X'_1\in \Delta(\der{D}'_1)$,
  $X'_2\in \Delta(\der{D}'_2)$, we denote their coprojections
  into, respectively, $\lpro \der{D}_1\rpro$, $\lpro \der{D}'_2\rpro$,
  $\lpro \der{D}'_1\rpro$, $\lpro \der{D}'_2\rpro$ by
  $\iota_{1, X_1}\colon X_1\to \lpro\der{D}_1\rpro$,
  $\iota_{1, X_2}\colon X_2\to \lpro\der{D}_1\rpro$,
  $\iota'_{1, X'_1}\colon X'_1\to \lpro\der{D}'_1\rpro$ and
  $\iota'_{2, X'_2}\colon X'_2\to \lpro\der{D}_1\rpro$. Similarly,
  given $X\in \Delta(\der{D})$ and $\X'\in \Delta(\der{D}')$ we
  use $\iota_X\colon X\to \tpro{D}$ and
  $\iota'_{X'}\colon X'\to \lpro \der{D}' \rpro$ for the
  coprojections.

  Now that these preliminaries matters have been addressed, we can proceed with the rest of the proof. We split the proof in three cases.
  \begin{itemize}
  \item $\lgh(\der{D}_1)=0$. Thus $\lgh(\der{D}'_1)=0$ too  and we have
    \begin{align*}
      (\der{D}, \alpha, \omega) & =(\der{D}_2, \alpha_2\circ \omega_1^{-1}\circ \alpha_1, \omega_2) \\ (\der{D}', \alpha', \omega')&=(\der{D}'_2, \alpha'_2\circ (\omega'_1)^{-1}\circ \alpha'_1, \omega'_2)
    \end{align*}
    Moreover, in this case $\sigma$ must be $\id{\emptyset}$ and $\sigma+\tau$ must be equal to $\tau$. The thesis now follows from the commutativity of the following diagram.
    \[\xymatrix@C=35pt{& G_{1,0}\ar[dd]^{\xi_{\id{\emptyset}}} \ar[dr]^{\omega^{-1}_1}&&G_{2,0}\ar[r]^-{\iota_{2, G_{2,0}}} & \lpro \der{D}_2\rpro \ar[dd]^{\xi_{\tau}}\\\pi(G_{1,0})  \ar[ur]^{\alpha_1} \ar[dr]_{\alpha'_1}&& \pi(G_{2,0}) \ar[ur]^{\alpha_2} \ar[dr]_{\alpha'_2}\\& G'_{1,0} \ar[ur]_{(\omega'_1)^{-1}}&&G'_{2,0} \ar[r]_-{\iota'_{2, G_{2,0}}}& \lpro \der{D}'_2\rpro}\]

@@

  If a switch of $\dder{D}_0$ and $\dder{D}_1$ exists we say that
  they are \emph{switchable}.

@@

\begin{notation} We  denote any switch of two switchable direct derivations $\dder{D}_0$ and $\dder{D}_1$ along $(i_0, i_1)$ by $S_{i_0, i_1}(\dder{D}_0, \dder{D}_1)$. Moreover, we use $S_{i_0,i_1}(\dder{D}_1)$ and  $S_{i_0,i_1}(\dder{D}_0)$ to denote its components, in such a way that \[S_{i_0, i_1}(\dder{D}_0, \dder{D}_1)=S_{i_0,i_1}(\dder{D}_1)\cdot S_{i_0,i_1}(\dder{D}_0)\]
\end{notation}

@@

%\begin{remark}
  %Clearly, a permutation $\perm$ is the identity if and only if
  %$\inv{\perm}=\emptyset$.
%\end{remark}

@@

%We also need to introduce the notion of \emph{switching sequences}.

@@

\begin{definition}[Switch equivalence]
  \label{de:switch-equivalence}
  Let $(\X, \R)$ be a left-linear rewriting system.  Let also
  $\der{D}, \der{E} : G \Mapsto H$ be derivations with the same
  length, $\der{D}=\{\dder{D}_{i}\}_{i=0}^n$ and
  $\der{E}=\{\dder{E}_{i}\}_{i=0}^n$. If
  $\dder{D}_i \cdot \dder{D}_{i+1}$ is a switch of
  $\dder{E}_i \cdot \dder{E}_{i+1}$ for some $i \in [0,n-1]$ `and  $\dder{D}_j = \dder{E}_j$ for each $j \not \in \{i,i+1\}$. We write
  $\der{D} \shift{\transp{i}} \der{E}$.

  A \emph{switching sequence} is  a sequence $\{\der{D}_{k}\}_{k=0}^m$ of derivations such that 
  \[\der{D}_0 \shift{\transp{i_1}} \der{D}_1 \shift{\transp{i_2}} \ldots \shift{\transp{i_m}}
  \der{D}_m\]
 In such case, we denote by $\nu_{k}$ the transposition $\transp{i_k}$. Given $k \leq h$,
 we denote by $\nu_{k,h}$ the composition
 $\nu_h \circ \nu_{h-1} \circ \ldots \nu_k$. We say that the
 switching sequence consists of inversions if for all
 $k \in \interval[0]{m}$ the transposition $\nu_k$ is an inversion for
 $\nu_{1,m}$.

  We say that two derivations $\der{D}, \der{E}:G\Mapsto H$ are \emph{switch
    equivalent} if there is a switching sequence, possibly empty, $\{\der{D}_{k}\}_{k=0}^m$ such that
      \[\der{D}\equiv_a \der{D}_0 \shift{\transp{i_1}} \der{D}_1 \shift{\transp{i_2}} \ldots \shift{\transp{i_m}}
    \der{D}_m\equiv _a \der{E}\]
    We denote this relation by $\der{D}\equiv^{sh} \der{E}$. 
\end{definition}

@@

  Consider the pullback of $g_0$ and $f_1$, getting $p_i : P \to D_i$
  for $i \in \{0,1\}$ and the mediating arrows $u_0\colon K_0\to P$ and
  $u_1\colon K_1\to P$ into the pullback object (see \Cref{prop:tec} for details). 
  % given by \Cref{prop:tec} .
  %
  We say that $(i_0, i_1)$ is a \emph{strong independence pair} if
  the first two squares depicted  
  below are pushouts and if the pushout of $r_1 : K_1 \to R_1$ and $u_1 : K_1 \to P$ exists
  \[
    \xymatrix{
      K_0 \ar[r]^{r_0} \ar[d]_{u_0}& R_0 \ar[d]^{i_0} & K_1
      \ar@{>->}[r]^{l_1} \ar[d]_{u_1}& L_1 \ar[d]^{i_1}
       &K_1 \ar[r]^{r_1} \ar[d]_{u_1}& R_1\ar@{.>}[d]^{j_0}
      \\
      %
      P \ar[r]_{p_1} & D_1 &P \ar[r]_{p_0} & D_0
       & P \ar@{.>}[r]_{q_1} & Q_1
    }
  \]
\end{definition}

@@

\begin{definition}[Graph signature]
  Let $\mathcal{G}=(V, E, s, t)$, with $s,t\colon E\rightrightarrows V$, be a finite directed acyclic graph,
  we define the \emph{depth} of a vertex $v\in V$ as
  \[
    \dph(v)=\begin{cases}
      0 & \text{there is no $e\in E$ such that $t(e)=v$}\\
      m+1 &\text{where } m=\min\{\dph(s(e)) \mid e\in E \text{ such that } t(e)=v
      \} 
    \end{cases}
  \]
\end{definition}

@@

As in \cite{BaldanGM06}, one can study  left-linear rewriting systems  on $\cat{eGraphs}$ in which the right half of a rule is mono but not regular in general, allowing us to identify, but not merging, more nodes . These rewring systems are well-switching.
\end{example}

@@

\subsection{Some results on $\mathcal{A}$-stable and $\mathcal{A}$-Van Kampen squares}
We start proving some general results regarding $\mathcal{A}$-Van Kampen and $\mathcal{A}$-stable squares. Let us begin recalling some classical results about  about pullbacks and pushouts\cite{mac2013categories}.

@@

In this appendix we provide the proofs of the main results of this paper.
